\newpage
\section{\textsc{State of the Art}}
\label{chap:soa}
\rhead{\itshape Chapter 1: State of the Art}
\lhead{}
\cfoot{\thepage}
\pagestyle{fancy}
\renewcommand{\headrulewidth}{0.4pt}

\begin{flushright}
\textit{``Research is to see what everybody else has seen,\\
and to think what nobody else has thought.''\\
Albert Szent-Györgyi}
\end{flushright}

\subsection{Automated Question Generation}

Early AQG systems applied rule-based syntactic transformations to produce
wh-questions \cite{heilman2010good}. Neural approaches fine-tuned
sequence-to-sequence models on SQuAD \cite{rajpurkar2016squad}, improving fluency
but lacking domain specificity and cognitive-level control.
The emergence of LLMs changed the paradigm: prompting alone (without fine-tuning)
now achieves competitive results.

\subsection{Reference Studies}

Four recent studies directly informed ExamGen's design.

\textbf{\citet{mucciaccia2025mcq} (COLING 2025)} process Portuguese university PDFs
through a multi-stage pipeline: PDF preprocessing, glossary extraction, question
generation, and an LLM-as-judge evaluation component consistent with human expert
judgements. Key contribution adopted in ExamGen: schema-enforced generation +
automated review process.
\textit{Limitation:} single language (Portuguese), no IDP or bilingual retrieval.

\textbf{\citet{isley2025quality}} conduct the largest field study of AI-generated
assessments (91 courses, $\approx$1,700 students). Using Item Response Theory,
they show AI-generated items perform comparably to expert-created ones when
tailored to course materials with iterative refinement.
Key contributions adopted: RAG grounding of generation; post-generation evaluator.
\textit{Limitation:} requires real student data for IRT validation.

\textbf{\citet{morris2024aqg}} introduce \textit{answer extractability} --- whether
the correct answer is derivable from the source material --- as a primary quality
metric, alongside structured grading rubrics for constructed-response questions.
Both concepts are directly implemented in ExamGen's evaluator and SAQ rubric.

\textbf{\citet{meissner2024itemforge}} validate Bloom's taxonomy-aligned generation
for mathematics, finding high cognitive-level alignment but occasional inaccuracies
in sample solutions. This motivates ExamGen's independent Bloom classifier (Stage~4)
rather than trusting the generator's self-reported level.

\subsection{Supporting Technologies}

\paragraph{Retrieval-Augmented Generation.}
\citet{lewis2020rag} propose RAG for knowledge-intensive NLP tasks. Hybrid
sparse-dense retrieval \cite{ma2021replication} with Reciprocal Rank Fusion
\cite{cormack2009rrf} improves over either approach alone. ExamGen extends this
with language-aware routing: AraBERT \cite{antoun2020arabert} for Arabic,
MiniLM \cite{reimers2019sentence} for French.

\paragraph{Intelligent Document Processing.}
Docling \cite{docling2024} provides TableFormer-based structure-preserving PDF
conversion with RTL-safe export. PyMuPDF \cite{pymupdf2023} serves as fallback
for formula recovery and code block extraction.

\subsection{Positioning of ExamGen}

\begin{table}[H]
\centering
\caption{ExamGen vs.\ reference systems.}
\label{tab:rw}
\resizebox{\textwidth}{!}{%
\begin{tabular}{lcccc}
\toprule
\textbf{Feature} & \textbf{Mucciaccia'25} & \textbf{Isley'25}
  & \textbf{Morris'24} & \textbf{ExamGen} \\
\midrule
Bilingual corpus              & \texttimes & \texttimes & \texttimes & \checkmark \\
IDP pipeline (5 stages)       & partial    & \texttimes & \texttimes & \checkmark \\
VLM diagram transcription     & \texttimes & \texttimes & \texttimes & \checkmark \\
Language-aware hybrid RAG     & \texttimes & \checkmark & \checkmark & \checkmark \\
Schema-enforced generation    & \checkmark & \checkmark & \texttimes & \checkmark \\
Independent Bloom classifier  & \texttimes & \texttimes & \texttimes & \checkmark \\
Post-generation evaluator     & \checkmark & \checkmark & \texttimes & \checkmark \\
Multi-tenant web application  & \texttimes & \texttimes & \texttimes & \checkmark \\
\bottomrule
\end{tabular}}
\end{table}

Table~\ref{tab:rw} reveals a clear gap: no existing system simultaneously addresses
bilingual IDP, language-aware hybrid retrieval, and an independent cognitive-level
evaluator.
ExamGen is designed precisely to fill this gap.
The design decisions required to do so --- iterative architecture, schema enforcement,
and Bloom classification --- are analysed in Chapter~\ref{chap:arch}.
